---
title: "Running logistic growth by cultivar"
author: "Jessica Guo"
date: "1/4/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Overview
This script applies a hierarchical Bayesian model for logistic growth to obtain 3 parameters for each cultivar: maximum and minimum height (cm), and maximum growth rate (cm/gdd). 


Load libraries
```{r}
library(dplyr)
source("source/fit_logistic_growth.R")
library(ggplot2)
library(forcats)
```

Read in MAC seasons 4 and 6 cleaned height data. 
```{r}
season4 <- na.omit(read.table(file = "data_clean/season4_combined.txt", sep = "\t",
                              header = TRUE,
                              stringsAsFactors = FALSE))

length(c(table(season4$cultivar)))
season6 <- na.omit(read.table(file = "data_clean/season6_combined.txt", sep = "\t",
                              header = TRUE,
                              stringsAsFactors = FALSE))
length(c(table(season6$cultivar)))
```

## Running models with lapply for all cultivars
Add sites as a numeric value for model run. Split each season into separate lists by cultivar. Check number of sites per cultivar. 
```{r}
# Function to turn site into numeric within cultivar
add_site <- function(x) {
  x$site <- as.numeric(as.factor(x$sitename))
  return(x)
}

# split into list of 336 or 326 cultivars and add sites
s4 <- subset(season4, select = c(sitename, gdd, canopy_height, 
                                 cultivar, date))[order(season4$date),] 
s4.list <- split(s4, f = s4$cultivar)
s4.list <- lapply(s4.list, FUN = add_site)

s6 <- subset(season6, select = c(sitename, gdd, canopy_height, 
                                 cultivar, date))[order(season6$date),] 
s6.list <- split(s6, f = s6$cultivar)
s6.list <- lapply(s6.list, FUN = add_site)

# Check sites per cultivar
nsite <- function(x) {length(unique(x$site))}
summary(do.call(rbind, lapply(s4.list, nsite))) # between 2 and 11
summary(do.call(rbind, lapply(s6.list, nsite))) # between 1 and 99, with 99 a distinct outlier

# Remove cultivar SP1516 at 99 sites
# Not enough observations per site to allow for model convergence with hierarchical model
rem.cult <- names(s6.list)[which.max(do.call(rbind, lapply(s6.list, nsite)))]
temp <- filter(s6, cultivar != rem.cult)
temp2 <- split(temp, f = temp$cultivar)
s6.list.clean <- lapply(temp2, FUN = add_site)

# Keep SP1516 for simple model
s6.sp1516 <- filter(s6, cultivar == rem.cult)
```

Run model for both seasons. 
```{r}
# run for season 4
start <- proc.time()
out.s4 <- lapply(s4.list, FUN = fit_logistic_growth, 
                        type = "RE", outdir = "model_output/season4")
end <- proc.time()
dur <- (end-start)[[3]]/60/60

# run for season 6
start <- proc.time()
# 325 cultivars
out.s6 <- lapply(s6.list.clean, FUN = fit_logistic_growth,
                 type = "RE", outdir = "model_output/season6")
# cultivar SP1516
fit_logistic_growth(data = s6.sp1516,
                    type = "simple",
                    outdir = "model_output/season6")
end <- proc.time()
dur <- (end-start)[[3]]/60/60
```

### Combine and plot

Read in cultivar-specific output. 
```{r}
# from saved model outputs
# Season 4
cultivars <- list.files("model_output/season4")
out.s4 <- list()
for (i in 1:length(cultivars)){
  load(file.path("model_output/season4", cultivars[i], "out.Rdata"))
  out <- out %>%
    mutate(season = "season_4")
  out.s4[[i]] <- out
}

# Season 6
cultivars <- list.files("model_output/season6")
out.s6 <- list()
for (i in 1:length(cultivars)){
  load(file.path("model_output/season6", cultivars[i], "out.Rdata"))
  out <- out %>%
    mutate(season = "season_6")
  out.s6[[i]] <- out
}

# combine into single dataframe
s4.df <- do.call(rbind, out.s4)
s6.df <- do.call(rbind, out.s6)
all.df <- rbind(s4.df, s6.df) %>%
  arrange(r2)

# Create output
out.csv <- all.df %>%
  select(season, cultivar, type,
         Ymax.median, # Ymax.lower, Ymax.upper,
         Ymin.median, # Ymin.lower, Ymin.upper,
         Ghalf.median, # Ghalf.lower, Ghalf.upper,
         r2) %>%
  rename(genotype = cultivar, method_type = type,
         max_height_cm = Ymax.median,
         min_height_cm = Ymin.median,
         max_growth_cm_gdd = Ghalf.median) %>%
  mutate(method = "Bayesian logistic model of canopy height vs. gdd") %>%
  relocate(method_type, .after = last_col()) %>%
  arrange(season,r2)

write.csv(out.csv, file = "data_clean/mac_growth_rate_modeled.csv", row.names = F)
```

Plot maximum height and growth rate by cultivar and season. 
```{r}
# Maximum height by cultivar and season
all.df %>%
  mutate(cultivar = fct_reorder(cultivar, Ymax.median)) %>%
  ggplot(aes(x = cultivar, y = Ymax.median)) + 
  geom_point(aes(color = season)) +
  # geom_errorbar(aes(ymin = Ymax.lower, ymax = Ymax.upper), alpha = 0.25) +
  scale_y_continuous("Ymax (cm)") +
  theme_bw(base_size = 12) +
  coord_flip()

# Max growth rate by cultivar and season
all.df %>%
  mutate(cultivar = fct_reorder(cultivar, Ghalf.median)) %>%
  ggplot(aes(x = cultivar, y = Ghalf.median)) + 
  geom_point(aes(color = season)) +
  # geom_errorbar(aes(ymin = Ghalf.lower, ymax = Ghalf.upper), alpha = 0.25) +
  scale_y_continuous("Ghalf (cm/gdd)") +
  theme_bw(base_size = 12) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  coord_flip()
```


