---
title: "Running logistic growth by cultivar"
author: "Jessica Guo"
date: "1/4/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Overview
This script applies a simple Bayesian model for logistic growth to season 6 canopy height data to obtain 3 parameters for each cultivar: maximum and minimum height (cm), and maximum growth rate (cm/gdd). 


Load libraries
```{r}
library(dplyr)
source("source/fit_logistic_growth.R")
library(ggplot2)
library(forcats)
```

Read in MAC 6 cleaned height data. 
```{r}
season6 <- na.omit(read.table(file = "data_clean/season6_combined.txt", sep = "\t",
                              header = TRUE,
                              stringsAsFactors = FALSE))
length(c(table(season6$cultivar)))
```

## Running models with lapply for all cultivars
Add sites as a numeric value for model run. Split each season into separate lists by cultivar. Check number of sites per cultivar. 

```{r}
# Function to turn site into numeric within cultivar
add_site <- function(x) {
  x$site <- as.numeric(as.factor(x$sitename))
  return(x)
}

# split into list 326 cultivars and add sites
s6 <- subset(season6, select = c(sitename, gdd, canopy_height, 
                                 cultivar, date))[order(season6$date),] 
s6.list <- split(s6, f = s6$cultivar)
s6.list <- lapply(s6.list, FUN = add_site)

# Check sites per cultivar
nsite <- function(x) {length(unique(x$site))}
summary(do.call(rbind, lapply(s6.list, nsite))) # between 1 and 99, with 99 a distinct outlier
```

Run model for season 6, using type = "simple" regardless of the number of locations. May have to restart for loop as needed. 
```{r}
# run for season 6
start <- proc.time()
# 326 cultivars

for(i in 1:length(s6.list)){
  fit_logistic_growth(s6.list[[i]], type = "simple",
                      outdir = "model_output/season6")
  print(i)
}
# out.s6 <- lapply(s6.list, FUN = fit_logistic_growth,
#                  type = "simple", outdir = "model_output/season6")
end <- proc.time()
dur <- (end-start)[[3]]/60/60
```

### Combine and plot

Read in cultivar-specific output and combine
```{r}
# Season 6
cultivars <- list.files("model_output/season6")
out.s6 <- list()
for (i in 1:length(cultivars)){
  load(file.path("model_output/season6", cultivars[i], "out.Rdata"))
  out <- out %>%
    mutate(season = "season_6")
  out.s6[[i]] <- out
}

# combine into single dataframe
s6.df <- do.call(rbind, out.s6)

```

Check whether model converged for all cultivars and compare Ymax and 
```{r}
summary(s6.df$Rhat)

s6.df %>%
  ggplot() +
  geom_point(aes(x = Ymax.median, y = Ghalf.median))

cor.test(~ Ghalf.median + Ymax.median, data = s6.df)
```


Output as csv. 
```{r}
# Create output
out.csv <- s6.df %>%
  select(season, cultivar, type,
         Ymax.median, # Ymax.lower, Ymax.upper,
         Ymin.median, # Ymin.lower, Ymin.upper,
         Ghalf.median, # Ghalf.lower, Ghalf.upper,
         r2) %>%
  rename(genotype = cultivar, method_type = type,
         max_height_cm = Ymax.median,
         min_height_cm = Ymin.median,
         max_growth_cm_gdd = Ghalf.median) %>%
  mutate(method = "Bayesian logistic model of canopy height vs. gdd") %>%
  relocate(method_type, .after = last_col()) %>%
  arrange(season,r2)

write.csv(out.csv, file = "data_clean/mac_growth_rate_modeled_season6.csv", row.names = F)
```

Plot maximum height and growth rate by cultivar and season. 
```{r}
# Maximum height by cultivar and season
s6.df %>%
  mutate(cultivar = fct_reorder(cultivar, Ymax.median)) %>%
  ggplot(aes(x = cultivar, y = Ymax.median)) + 
  geom_point(aes(color = season)) +
  # geom_errorbar(aes(ymin = Ymax.lower, ymax = Ymax.upper), alpha = 0.25) +
  scale_y_continuous("Ymax (cm)") +
  theme_bw(base_size = 12) +
  coord_flip()

# Max growth rate by cultivar and season
s6.df %>%
  mutate(cultivar = fct_reorder(cultivar, Ghalf.median)) %>%
  ggplot(aes(x = cultivar, y = Ghalf.median)) + 
  geom_point(aes(color = season)) +
  # geom_errorbar(aes(ymin = Ghalf.lower, ymax = Ghalf.upper), alpha = 0.25) +
  scale_y_continuous("Ghalf (cm/gdd)") +
  theme_bw(base_size = 12) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  coord_flip()
```


